---
title: "R Notebook"
output: html_notebook
---


First load the required packages


```{r}
require(WDI)
require(countrycode)
require(dplyr)
require(googleVis)
```


let's first read in the data


Let's set the indicator names and corresponding variable names.

before setting indicator names and others I should mention a little about how do I find the indicator names. For example I was looking for gdp per capita in constant dollars. so i first employed the following 

```{r}
WDIsearch("gdp per capita") %>% dim()
```

Let's have a quick check on the class

```{r}
WDIsearch("gdp per capita") %>% class()
```

Actually we could assign the following command

```{r}
str(WDIsearch("gdp per capita"))
```

It gives us a little bit more information. We see `WDIsearch()` funciton basically returns a matrix of 2 columns and 13 rows

In that case we could just browse and find the corresponding indicator name. but in case there is large number of resutls, we could autmate the process a bit. It may come useful because sometimes it becomes difficult to find the corresponding names in case the results are quite large. so here we go

first save the search result

```{r}
wdi_sch <- WDIsearch("gdp per capita")
```

Find the row index of our intended indicator

```{r}
grep("GDP per capita (.+) 2000 (.+)",wdi_sch[,2],value =T,ignore.case = T)
```

in the above we see that we were able to find the right kind of result

so we now save the indicator index

```{r}
indic <- grep("GDP per capita (.+) 2000",wdi_sch[,2],ignore.case = T)
```

so i got the row indicator, then i find it from the first column

```{r}
wdi_sch[indic,1]
```


Now let's set this in the list of indicators 

```{r}
inds <- c('SP.DYN.TFRT.IN','SP.DYN.LE00.IN', 'SP.POP.TOTL',
'NY.GDP.PCAP.KD', 'SE.ADT.1524.LT.FE.ZS', 'SH.DYN.MORT')
indnams <- c("fertility_rate", "life_expectancy", "population",
"GDP_per_capita", "female_literacy","mortality_rate")
```

Time to get the country code. question is what are the countries to choose? my reference point was name of the countries in this article 

http://www.forbes.com/sites/timworstall/2016/11/26/fidel-castros-economic-disaster-in-cuba/

We will use countrycode package to get required codes

```{r}
countries <- countrycode(c("Cuba","Ecuador","Jamaica","Panama"),'country.name','iso2c')
```

Just have a quick look at the countries

```{r}
countries
```

Looks just fine.


Now let's get the data

```{r}
wdi_all <- WDI(country='all', 
              indicator=inds,
              start=1960, 
              end=2016, 
             extra=TRUE)
```


Now in the following, we will be replacing the column headers with the indicator names. Othewise, the names are the indicators which are difficult to undersand. 

```{r}
colnum <- match(inds, names(wdi_all))
names(wdi_all)[colnum] <- indnams
var_names <- gsub("\\.|\\s","_",names(wdi_all))
var_names <- sub("15.to.25.yr.female.literacy","female_literacy",var_names)
names(wdi_all) <- var_names
```

Time to save the file
```{r}
save(wdi_all,file="./RDATA/wdi_all.RData")

```

Now let's have a look at the regions

```{r}
wdi_all %>% group_by(region) %>%
  summarise(num=n())
```

in the above we see a category with NA, let's check that out

```{r}
wdi_all %>% group_by(income) %>%
  summarise(num=n())
```

let's check out what are those NAs

```{r}
wdi_all %>% 
  filter(is.na(income)) %>%
  head()
```

now let's take care of latin america

```{r}
ex_vars = c("iso3c","region","capital","longitude","latitude","income","lending")  

in_vars <- c("iso2c", "country","year", "fertility_rate","life_expectancy"            ,"population" , "GDP_per_capita_Constant_USD", "female_literacy", "Mortality_Rate"  ,"iso3c","region")

in_vars_list <- as.list(in_vars)

```

         

```{r}
latin <- wdi_all %>%
  filter(region=="Latin America & Caribbean (all income levels)") %>%
  select_(.dots=in_vars_list)

save(latin,file="./RDATA/latin.RData")
```


# Now time for the motion chart

```{r, results='asis',tidy=FALSE}
myState <- '{"yZoomedIn":false,"xZoomedDataMax":81.49619512,"yLambda":1,"iconType":"BUBBLE","xLambda":1,"xZoomedDataMin":42.10009756,"yZoomedDataMin":1.47,"nonSelectedAlpha":0.4,"orderedByX":false,"xAxisOption":"4","sizeOption":"5","uniColorForNonSelected":false,"orderedByY":false,"yZoomedDataMax":7.555,"dimensions":{"iconDimensions":["dim0"]},"showTrails":true,"colorOption":"2","time":"2014","duration":{"timeUnit":"Y","multiplier":1},"playDuration":15000,"yAxisOption":"3","iconKeySettings":[{"key":{"dim0":"Cuba"},"trailStart":"1960"}],"xZoomedIn":false}'

myState <- '{"iconKeySettings":[{"key":{"dim0":"Cuba"},"trailStart":"1960"}]}'

M <- gvisMotionChart(latin,
                      idvar="country", timevar="year",
                      xvar="life_expectancy", yvar="fertility_rate",
                      colorvar="iso2c", sizevar="population",
                      options=list(width=640,
                                   #state=myState,
                                   height=420))
## Display the chart in the browser
plot(M)
```

The problem with the above is that when we tried to impose  `state` variable, it isn't working. 

Let me try single variables, just the life expectancy


```{r}
LE_MR<- gvisMotionChart(latin,
                      idvar="country", timevar="year",
                      xvar="life_expectancy",yvar = "Mortality_Rate",
                      colorvar="iso2c", sizevar="population",
                      options=list(width=640,
                                   #state=myState,
                                   height=420))
## Display the chart in the browser
plot(LE_MR)
```

